Cartogram Mapping and its Application to Cancer Data Visualization

### 2019-12-09 Response/Response.Rmd

---
from: Stephanie Kobakian
fromqualifications: BA
fromtitle: Research Masters Student
fromemail: stephanie.kobakian@monash.edu
fromphone: +61 433 699 797
fromsignature: sigfile.png
toname: Professor Peter Baade and Dr Susanna Cramb
opening: Dear Professor Peter Baade and Dr Susanna Cramb,
closing: Warm regards
spacing: 1.4
output: MonashEBSTemplates::letter
---

@@

**Editor suggestion:**
**We would suggest that the aim of the manuscript could be phrased to highlight the current approaches and limitations of typical visualisation methods for cancer maps, along with outlining possible alternative methods and challenges in implementing these?**

We have restructed our original submission to provide a clear path from current approaches used, to the alternative approaches that should be explored. We highlight the limitations of using current display methods, to encourage the use of alternative display methods. We also present the challenges that may be faced when implementing the alternative designs.

The manuscript now contains the following sections:

**Traditional approaches for cancer map displays:** This section discusses the choropleth map as the common map display used in cancer maps and atlases.
*Limitations of current approaches:* This section presents the possible issues with displaying cancer statistics on a choropleth map.
*Alternative visualisation methods:* We have shown that there are many alternative displays that highlight population dense regions.
*Challenges of alternative approaches:* Alternative displays also have limitations when displaying spatial distributions of cancer statistics.

@@

We have inclued changes to reflect your minor notes:

*Editor suggestion:*
*The authors mention a focus on Atlases published between 2010 and 2015, but then include a bowel cancer atlas from 2016. *

*Also, within the bowel cancer atlas data, the reporting of percent of males aged 50-54 with bowel cancer in 2016 is incorrect; it should be the % of the ERP in each area that are males aged 50-54 years. *

We have changed the image used from the Bowel Cancer Atlas. To reflect this change the description of the statistic used in the bowel cancer atlas of Australia ahs been corrected:

Bowel Cancer Australia Atlas:
...shows the average Standardised Incidence Ratio of colorectal cancer for Australian males from 2006 to 2010 in Australia. It is published by Bowel Cancer Australia https://www.bowelcanceraustralia.org/.

*Reviewer A: *
*Major*
*Comment 1: *
The Abstract should be more focused- perhaps give an example of how cartogram mapping has been applied to data visualisation. What are the benefits of this approach and suggestions for the future.

*Reply 1:* 
We agree with the reviewer. Alternative mapping displays have been implemented in political contexts. Choropleth maps have often been used.

### 2019-12-09 review_article.Rmd
response letter

<!-- Structure of the paper -->
The paper is organized as follows. The next section describes the choropleth map which is the common approach to disease maps and presents examples of atlases in use today. The limitations of the choropleth map are discussed in Section \ref{ch:chorolimit}. Section \ref{ch:alternatives} describes alternative displays, including the cartogram which is useful when the map has heterogeneously sized geographic units. Section \ref{ch:altlimits} presents the limitaions of producing and using alternative displays. Disease maps are more useful when made interactive, and common options are described in Section \ref{ch:interacting}, along with a discussion of benefits and disadvantages.

@@

## Overview of publicly available atlases \label{sec:publicatlases}
<!-- 416 words -->
Roberts [@roberts2019communication] identified 33 publicly available cancer atlases, published between 2010 and 2018, from different countries around the globe. All of these online atlases used choropleth maps. All but one of these were published by non-commercial organizations, including not-for-profits, government, research organizations, advocacy groups or government-funded partnerships.  Figure \ref{fig:choropleth-grid-create} displays a subset of maps from these atlases.

@@

## Additional considerations
<!-- 479 words -->
Maps cannot tell the entire story. Supplementary graphs and plots are often included to add more depth and information. Additional materials such as tables, graphs, and text explanations, support understanding and inference derived from maps, ensuring the message communicated will be consistent across a range of viewers [@CPISACA]. The many displays of statistical summaries, including dot plots, bar plots, box plots, cumulative distribution plots, scatter plots, normal probability plots, can provide alternative views of the cancer statistics. These can also display supporting statistics such as error, confidence intervals, distributions, sample or population sizes, standard deviation. When presenting cancer maps, the intuition derived from maps must be 'validated by rigorous statistical analyses' [@MACM], and the display of the supplementary statistics helps in this regard.

@@

 As this feature of population distributions continues to intensify, the need for cartograms as an alternative to a choropleth map should only increase.
# Contemporary alternatives to choropleths {#ch:alternatives}

@@

# Limitations of alternative displays {#ch:compare}
<!-- 408 words -->

@@

Many publicly available cancer atlases employ interactivity and/or animation. Figure 6 <!--\ref{fig:interacting}--> shows the interactive GUI tools provided with the public atlases discussed in Section \ref{sec:publicatlases}. Mostly these are menus, and toggles, for choosing indicators, and subsets of the population.

>>>>

<!-- Structure of the paper -->
The paper is organized as follows. The next section describes the choropleth map which is the common approach to disease maps and presents examples of atlases in use today. The limitations of the choropleth map are discussed in Section \ref{ch:chorolimit}. Section \ref{ch:alternatives} describes alternative displays, including the cartogram which is useful when the map has heterogeneously sized geographic units. Section \ref{sec:altlimits} presents the limitaions of producing and using alternative displays. Disease maps are more useful when made interactive, and common options are described in Section \ref{ch:interacting}, along with a discussion of benefits and disadvantages.

@@

<!-- 416 words -->
Roberts [@roberts2019communication] identified 33 publicly available cancer atlases, published between 2010 and 2018, from different countries around the globe. All of these online atlases used choropleth maps. All but one of these were published by non-commercial organizations, including not-for-profits, government, research organizations, advocacy groups or government-funded partnerships.  Figure \ref{fig:choropleth-grid-create} displays a subset of maps from these atlases.

@@

## Additional considerations
<!-- 479 words -->
Supplementary graphs and plots are often included to add more depth and information. Additional materials such as tables, graphs, and text explanations, support understanding and inference derived from maps, ensuring the message communicated will be consistent across a range of viewers [@CPISACA]. The many displays of statistical summaries, including dot plots, bar plots, box plots, cumulative distribution plots, scatter plots, normal probability plots, can provide alternative views of the cancer statistics. These can also display supporting statistics such as error, confidence intervals, distributions, sample or population sizes, standard deviation. When presenting cancer maps, the intuition derived from maps must be 'validated by rigorous statistical analyses' [@MACM], and the display of the supplementary statistics helps in this regard.

@@

 As this feature of population distributions continues to intensify, the need for cartograms as an alternative to a choropleth map should only increase.

# Contemporary alternatives to choropleths {#ch:alternatives}

@@

# Limitations of alternative displays {#sec:altlimits}
<!-- 408 words -->

@@

Many publicly available cancer atlases employ interactivity and/or animation. Figure 6 <!--\ref{fig:interacting}--> shows the interactive GUI tools provided with the public atlases. Mostly these are menus, and toggles, for choosing indicators, and subsets of the population.
